---
title: "Figshare workflow"
auther: Ernest Guevarra
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Figshare workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(pactr)
```

## About the Pandemic PACT Figshare repository

The Pandemic PACT data is meant to be publicly available for download from Figshare as per information from the research programme's website (see [here](https://www.pandemicpact.org/about/our-data)) and based on its open peer review paper on Wellcome Open Research describing the [study protocol](https://wellcomeopenresearch.org/articles/9-156/v1?src=rss).

However, the link provided from the website ([https://figshare.com/s/58527668245cb63f14f5](https://figshare.com/s/58527668245cb63f14f5)) leads to a privately-shared Figshare collection. This means that the collection is only accessible to those who created it and to those who have been authorised to access it. The contents of the collection can be seen. However, the individual items cannot be viewed interactively from the browser and cannot be downloaded. This was the case as at 24 August 2024 when development of this package was initiated and continue to current date. This is contrary to what is mentioned in the research programme's website:

> "The data set can be downloaded from this website and all metadata and the dataset can be downloaded from Figshare."

and in the study protocol under open peer review:

> "Finally, the database is saved as a csv file and made available to the public on Figshare and can additionally be downloaded from the online funding tracker."

Thinking that maybe this was just a mistake (i.e., that there is a private and public repository and the link to the private repository was the one that was mistakenly shared), a simple search was performed within Figshare using the keyword "Pandemic PACT" in the search field. This search yielded several results the most relevant of which was the following:

[https://portal.sds.ox.ac.uk/pandemicpact](https://portal.sds.ox.ac.uk/pandemicpact)

which is a University of Oxford-owned instance of Figshare containing the various data products (tracker data, data dictionary, scoping review data, extended data) referred to and used by the research programme. More importantly, this collection is public hence any individual, whether logged into Figshare or not, can view and download the datasets. Without further examining, it would be easy to think that this might be the publicly available data and data products being referred to by the website and the research protocol.

However, on closer inspection, the datasets within this collection were last updated in 2023 and the names of the file indicated that these were demonstration datasets.

Knowing that the research programme was indeed using Figshare to host its dataset (but not accessible) and given the existence of this demonstration dataset within the University of Oxford Figshare instance, development of Figshare-specific functions within the `{pactr}` package utilised the demonstration collection to prototype the Figshare-interface functions with the hope that if Pandemic PACT does decide to share their official dataset on Figshare publicly, the refactoring needed to get the current functions working will be minimal.

As such, the following guidance on use of the `{pactr}` Figshare interface functions should be used only for demonstration, development, and testing purposes and should not be used for production or research tasks.

## Figshare interface

This package provides an application programming interface (API) to the research programme's Figshare repository to provide programmatic access to its publicly available tracker data along with it's other data products.

The functions for Figshare interface are wrappers to specific functions of the [`{deposits}` package](https://docs.ropensci.org/deposits/index.html) which provides the underlying universal interface to various online research data deposition services including Figshare. Current Figshare-specific functionalities available in `{pactr}` are:

1. Listing of outputs/assets available from the Panedmic PACT Figshare repository (*experimental*);

2. Downloading of outputs/assets available from the Pandemic PACT Figshare repository (*experimental*);

3. Reading of dataset outputs/assets available from the Pandemic PACT Figshare repository (*experimental*); and,

4. Processing of Pandemic PACT data (*experimental*).

## Figshare workflow

The following diagram illustrates the current steps (along with their corresponding functions) in the designed API for interfacing with the Figshare repository.

```{r figshare-workflow, echo = FALSE, fig.align = "center"}
text <- paste(
  readLines("figshare.mmd"),
  collapse = "\n"
)

DiagrammeR::mermaid(text)
```

### Set a Figshare client

Usage of `{pactr}` always starts with the setting up of a Figshare client. This requires creating a Figshare account and then creating a personal access token [here](https://figshare.com/account/applications).

Once a Figshare token is created, it needs to be stored as a local environment variable. This can be done using the following command in R:

```R
Sys.setenv("FIGSHARE_TOKEN"="YOUR_TOKEN_HERE")
```

Once this token has been set as described above, the following command can be run to setup a Figshare client:

```{r setup-client}
pact_client <- pact_client_set()
```

Once a Figshare client has been setup, you can now perform the functionalities provided by the `{pactr}` package.

### List outputs/assets

To list available outputs/assets from the Pandemic PACT Figshare repository, issue the following command:

```{r usage-1a, eval = FALSE}
pact_list(pact_client)
```

The output is a `data.frame` containing metadata regarding contents of the Figshare Pandemic PACT group. The information within the metadata are those provided by [Figshare's application programming interface (API)](https://docs.figshare.com/) and are either set by the Pandemic PACT data team or by Figshare. The data.frame would look as follows:

```{r usage-1b, echo = FALSE}
pact_list(pact_client) |>
  tibble::tibble()
```

This function is useful in getting an overview of what is currently available in the Pandemic PACT Figshare repository.

### Download outputs/assets

To download a specific output/asset - say the scoping review data - from the Pandemic PACT Figshare repository, issue the following commands:

```{r usage-2a, eval = FALSE}
## Get the unique identifier for the scoping review data from Figshare ----
file_id <- pact_list(pact_client) |>
  subset(title == "Scoping Review Data", select = id) |>
  unlist()

pact_download_figshare(pact_client, id = file_id, path = ".")
```

This will download the file `Scoping_Review-Data.xlsx` from the Pandemic PACT Figshare repository into the current working directory.

### Read the Pandemic PACT tracker dataset and data dictionary

To read the Pandemic PACT tracker dataset into R, issue the following command:

```{r usage-3a, eval = FALSE}
pact_read_figshare(pact_client)
```

which outputs a data.frame of the available Pandemic PACT funding tracker dataset that has been labelled.

This function reads the *labelled* tracker dataset by default. If the raw dataset is required, then issue the following command:

```{r usage-3c, eval = FALSE}
pact_read_figshare(pact_client, tracker_type = "raw")
```

which outputs a data.frame of the available raw Pandemic PACT funding tracker dataset.

### Process the Pandemic PACT tracker dataset

```{r useage-3e, eval = FALSE}
pact_read_figshare(pact_client) |>
  pact_process_figshare()
```
